---
title: "How Can a Wellness Technology Company Play It Smart?"
author: "Adedotun Charles Adegunju"
date: "1/17/2022"
output: html_document
---

# Step 1: Ask
## Business Task
Gaining insights into how people are using smart devices to provide high-level recommendations for how these trends can inform marketing strategy.

# Step 2: Prepare
## Data Source
This dataset generated by respondents to a distributed survey via Amazon Mechanical Turk between 03.12.2016-05.12.2016. Thirty eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. Individual reports can be parsed by export session ID (column A) or timestamp (column B). Variation between output represents use of different types of Fitbit trackers and individual tracking behaviors / preference.
### Acknowlegement
Furberg, Robert; Brinton, Julia; Keating, Michael ; Ortiz, Alexa
https://zenodo.org/record/53894#.YMoUpnVKiP9

# Step 3: Process
## Installing and loading packages and libraries
```{r load packages and libraries, include=FALSE}
library(tidyverse)
library(lubridate)
```
## Loading data
```{r read data}
daily_activity <- read.csv("../data/dailyActivity_merged.csv")
daily_calories <- read.csv("../data/dailyCalories_merged.csv")
daily_intensities <- read.csv("../data/dailyIntensities_merged.csv")
daily_steps <- read.csv("../data/dailySteps_merged.csv")
heartrate_seconds <- read.csv("../data/heartrate_seconds_merged.csv")
hourly_calories <- read.csv("../data/hourlyCalories_merged.csv")
hourly_intensities <- read.csv("../data/hourlyIntensities_merged.csv")
hourly_steps <- read.csv("../data/hourlySteps_merged.csv")
minute_calories_narrow <- read.csv("../data/minuteCaloriesNarrow_merged.csv")
minute_calories_wide <- read.csv("../data/minutecaloriesWide_merged.csv")
minute_instensities_narrow <- read.csv("../data/minuteIntensitiesNarrow_merged.csv")
minute_instensities_nwide <- read.csv("../data/MinuteIntensitiesWide_merged.csv")
minute_mets_narrow <- read.csv("../data/minuteMETsNarrow_merged.csv")
minute_sleep <- read.csv("../data/minuteSleep_merged.csv")
minute_steps_narrow <- read.csv("../data/minuteStepsNarrow_merged.csv")
minute_steps_wide <- read.csv("../data/minuteStepsWide_merged.csv")
sleep_day <- read.csv("../data/sleepDay_merged.csv")
weight_log_info <- read.csv("../data/weightLogInfo_merged.csv")
```
## Exploring a few key tables
```{r}
head(daily_activity)
```
```{r}
colnames(daily_activity)
```
